Evaluation of pilot jobs for Apache Spark\\ applications on HPC clusters

### 2019-05-20 paper.tex
Edits and comments

\documentclass{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{ulem} % for \uwave
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\usepackage{stmaryrd}
\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}

@@

\begin{abstract}
    Big Data has become prominent throughout many scientific fields and, as
    a result, scientific communities have sought out Big Data frameworks to
    accelerate the processing of their increasingly data-intensive
    pipelines. However, while scientific communities typically rely on
    High-Performance Computing (HPC) clusters for the parallelization of
    their pipelines, many popular Big Data frameworks such as Hadoop and
    Apache Spark were primarily designed to be executed on dedicated
    commodity infrastructures. This paper evaluates the benefits of pilot
    jobs over traditional batch submission on HPC clusters with overlay
    Apache Spark clusters. We first evaluate the overall speedup brought on
    by employing pilot jobs. We then examine the robustness to master and
    driver failures, which may be frequent when underestimating pilot
    walltime. Spark built-in checkpointing is also investigated in relation
    to pilot expiration. Surprisingly, our results show that the speed-up
    provided by pilot jobs over batch scheduling remains moderate (\TG{X} on
    average) despite the presence of long queuing times. In addition, pilot
    jobs provide an extra layer of scheduling that complexifies debugging
    and deployment. We conclude that traditional batch scheduling should
    remain the default strategies to deploy Apache Spark applications on
    HPC clusters.

@@

Pilot jobs, also known as dynamic resource provisioning or glide-in
scheduling, are a widely-used technique to address infrastructure
heterogeneity, variable task queuing times, fine task granularity, and node
failures on distributed computing infrastructures~\cite{jakubsmodel}. Made popular by software
projects such as Condor~\cite{thain2005distributed} and DIRAC~\cite{casajus2010dirac}, they
became critical to grid computing, greatly improved the performance of HPC
clusters, and enabled multi-cloud executions.

@@

% Goal of the paper
In this paper, we study the use of pilot jobs to run scientific
applications with Apache Spark~\cite{zaharia2016apache} on shared HPC
clusters. The method currently recommended for this purpose involves batch
requesting all the
necessary resources and launching a standalone Spark cluster once the
resources have been allocated. This is for instance the method used in
Compute Canada (see details \href{https://docs.computecanada.ca/wiki/Apache_Spark/en}{here}),
 our national computing infrastructure. With pilot
jobs, rather than requesting all the resources at once, a Spark cluster is
launched with a subset of the resources, and is expanded as more resources
get allocated. Consistently with previous examples of pilot job
deployments, we hypothesize that this strategy would reduce queuing times
by (1) fragmenting resource requirements, and (2) allowing shorter walltime
estimates. While there have been some efforts on implementing pilot jobs
for Apache Spark~\cite{jha-spark-pbs}, research is limited and
none of the them detail their quantitative effect.

Apache Spark is a popular Big Data framework, commonly used in both
industrial and academic settings. Although it is a Scala-based framework,
it also has APIs for Java, Python (PySpark) and R. Spark's Resilient
Distributed Dataset (RDD) abstraction enabled in-memory processing of
pipelines by co-locating tasks and data, which provided important
performance improvements compared to its predecessor Hadoop
MapReduce~\cite{dean2008mapreduce}. Through the use of RDDs, it also became possible
to execute iterative workflows -- something not easily doable in older
frameworks. Schedulers for Spark include its built-in standalone schedule,
Yet Another Resource Negotiator (YARN~\cite{apache13yet}), and Mesos~\cite{hindman2011mesos}. As a
result of the sustained increase in data volumes, Apache Spark and other
Big Data engines are increasingly used for scientific applications,
including in neuroimaging, our primary field of
interest~\cite{boubela2016big,mehta2017comparative,maybethesimulationone,freeman2014mapping}.

@@

    \subsection{Infrastructure}
    All experiments were conducted on the Cedar and B\'eluga HPC computing
    clusters made available by \href{https://www.computecanada.ca}{Compute
    Canada} through \href{https://www.westgrid.ca}{WestGrid} and
    \href{http://www.calculquebec.ca}{Calcul Qu\'ebec}. Both clusters are
    accessible through the Slurm batch scheduler and Lustre parallel file
    system~\cite{schwan2003lustre}. The Cedar cluster has a total of 1542
    nodes with a total of 58,416 CPU cores. Available memory on a Cedar
    node can range from 125 to 3022~G. Standard nodes are equipped with
    either 2x Intel E5-2683 v4 Broadwell @ 2.1~Ghz (32 cores total) or 2x
    Intel Platinum 8160F Skylake @ 2.1Ghz (48 cores total) CPUs and 2 x
    480~G SSD. All nodes and temporary storage on Cedar are connected by an
    Intel OmniPath (version 1) with 100Gbit/s bandwidth.

    B\'eluga, on the other hand, is a smaller cluster with 872 available
    nodes. Node memory can range between 92 to 752~G, with the most common
    node type having 186G. All nodes contain 2 x Intel Gold 6148 Skylake @
    2.4~Ghz (40 cores/node) CPU and are connected to each other with a
    100~Gb/s Mellanox Infiniband EDR network. Each non-GPU
    node type contains one 480~G SSD.

    It is important to note that these clusters were used in production,
    that is, concurrently with other users. This allowed us to test the
    added-value of pilot scheduling in different realistic conditions of queuing
    times. At the time of our experiments, B\'eluga had recently been
    commissioned, resulting in low usage and shorter queuing times overall.
    Conversely, Cedar had been operating for a few years, resulting in
    higher usage and longer queuing times. Our experiments were repeated 15
    times in each configuration to capture queuing time variability.

@@

    A Spark cluster is made up of three components: the cluster manager, workers and
    driver. The cluster manager, also known as the Spark master in Standalone mode, is responsible for the
    resource provisioning within the cluster. Unlike the workers and driver, it is not
    application specific. The driver, which can start in the client where it is launched (i.e. client deploy mode)
    or directly on a worker node (i.e. cluster deploy mode), must request resources from the master
    to launch its tasks. The master the requests workers with the necessary resources available to start
    executors. The driver can then communicate with the executors to schedule and launch its tasks. Both the driver and
    the executors are application specific and connected directly to the master.

@@

    There is not shortage of scripts available online for starting a Spark cluster on HPC~\footnote{\url{https://github.com/NIH-HPC/spark-slurm},}\footnote{\url{https://www.sherlock.stanford.edu/docs/software/using/spark/},}\footnote{\url{https://sparkhpc.readthedocs.io},}\footnote{\url{https://www.osc.edu/~troy/pbstools/man/pbs-spark-submit}}.
    However, all these scripts start up a cluster within a single batch call. At most, there are two calls, one to startup the cluster,
    and the second, to start the application execution. None of these use any kind of pilot-scheduling approach, nor do they discuss the
    effects of queuing time and how they might want to adjust it in the case where the driver is started separately
    from the batch job to start the cluster. Even in the case of RADICAL-Pilot, as per some available tutorials for using the pilot-job application with
    Spark~\footnote{\url{https://github.com/radical-cybertools/pilot-streaming/blob/master/examples/Pilot-Streaming-GettingStarted.ipynb},}\footnote{\url{https://github.com/radical-cybertools/MIDAS-tutorial/blob/master/pilot/Pilot-Spark.ipynb}}, it appears that the Spark cluster is started within a single pilot and the application is submitted, in a separate process, to that pilot.
    Due to the fact that existing solutions mainly rely on a single batch call to startup the Spark
    cluster and do not appear to be able to submit multiple smaller jobs to improve queuing times, we have decided to implement our own solution.

    Two different job templates were used to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm, \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: certain resource requirements are requested by
    the user (e.g. walltime, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \texttt{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy mode selected for the batch template is client mode. We selected this
    deploy mode over cluster mode as driver recovery would not be required in batch in the
    case of walltime expiration, as all the allocated resources will expire at the same time.
    Moreover, many scientific applications are written in Python, which cannot use cluster mode using
    the Standalone resource manager. Should pilot scheduling using cluster mode significantly reduce
    queuing time compared to batch scheduling using client mode, it would provide enough justification to extend
    Spark's Standalone scheduler to include cluster mode for PySpark applications.

@@

    Both of these Slurm templates are launched within a Python application called
    \textit{SPA}. The templates are used in conjunction with JSON configuration 
    file and passed to the SlurmPy~\cite{slurmpy} library within \textit{SPA}.
    The \textit{SPA} application, all the while ensures that all is preconfigured
    correctly before passing it to SlurmPy. It also ensures that enough pilots are
    launched, maintains track of the running/queued pilots, and launches additional
    pilots if there are less pilots than requested by the user in the Slurm queue.

    \subsection{Application}
        \begin{algorithm}\caption{Incrementation}\label{alg:incrementation}              
            \begin{algorithmic}[1]                                                       
                \Input                                                                       
                    \Desc{$x$}{a sleep delay in seconds}                                         
                    \Desc{$n$}{a number of iterations}                                           
                    \Desc{$C$}{a set of image chunks}                                            
                \EndInput                                                                    
                \ForEach{$chunk$ in $C$}                                                      
                    \State read $chunk$                                        
                    \For{$i \in \llbracket 1, n \rrbracket$}                                                         
                        \State $chunk\gets chunk+1$                                              
                        \State sleep $x$                               
                    \EndFor                                                                      
                    \State write $chunk$                                            
                \EndFor                                                                      
            \end{algorithmic}                                                                
        \end{algorithm}
    To determine the added value of pilot scheduling over batch scheduling of Spark
    applications, we required a Spark application operating on a large dataset with
    an important processing time to emulate what would be the average requirements
    of a scientific Spark application. For this, we created a synthetic application 
    that would process the Big Brain~\cite{amunts2013bigbrain}, a 76~G 3D histological
    image of a human brain. The algorithm is a chain of map transformations that
    at each transformation increment the voxels of the image by 1 (see Algorithm~\ref{alg:incrementation}).
    We chose such a synthetic algorithm
    as the focus of our experiments is pilot scheduling and not the application in
    itself. Furthermore, this algorithm enabled us to have control over the task duration
    which was representative of scientific applications. Additionally, it was important
    that the overall application duration did not vary between the different levels of
    parallelism within our experiments. Being able to adjust the task duration based on
    level of parallelism allowed us to achieve this.

    Investigating the fault-tolerance provided by the overlay cluster is
    crucial in determining the suitability of pilot scheduling with overlay clusters.
    Fault-tolerance of the driver is only possible in cluster deploy mode, however,
    when using Spark's standalone scheduler, this mode is not available for Python 
    applications. It is for this reason that our synthetic application is written
    in Scala. Nevertheless, cluster deploy mode is possible for Python application
    using YARN or Mesos schedulers.

@@

    Assuming a divisible load, i.e., the application can be divided in any
    number of tasks, the makespan can be written using the following
    expression, which holds for both batch and pilot execution modes:
    \begin{equation}
        M = \frac{C}{W} \label{eq:mcw}
    \end{equation}
    where:
    \begin{itemize}
        \item $M$ is the makespan of the application
        \item $C$ is the total CPU time of the application
        \item $W$ is the average number of Spark workers throughout the execution
    \end{itemize}
    The average number of workers $W$ allows the model to take into account
     variable queuing times. It is computed as follows:
    \begin{equation}
        W = \frac{1}{M}\int_0^M{w(t)dt}\label{eq:avgw}
    \end{equation}
    where $w(t)$ is the number of workers available at time $t$. When the
    application is not subject to any scheduling or queuing time, the
    average number of workers equals the number of workers requested.

@@

        \begin{table*}                                                                   
        \centering                                                                       
        \begin{tabular}{c|cccccc}                                                   
          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 1}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 1 & 112 & 1 & 16 & 2h30 & 45 \\
          8 pilots & 1 & 14 & 1 & 2 & 2h30 & 45 \\
          16 pilots & 1 & 7 & 1 & 1 & 2h30 & 45 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 2}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 2 & 112 & 1 & 16 & 2h30 & 90 \\
          8 pilots & 1 & 28 & 1 & 4 & 2h30 & 90 \\
          16 pilots & 1 & 14 & 1 & 2 & 2h30 & 90 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 3}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 3 & 112 & 1 & 16 & 2h30 & 120 \\
          8 pilots & 1 & 42 & 1 & 6 & 2h30 & 120 \\
          16 pilots & 1 & 21 & 1 & 3 & 2h30 & 120 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 4}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          execution mode & nodes/job & RAM (G) & CPUs per task & tasks/node & walltime & task delay (s) \\                             
          \hline
          Batch & 4 & 112 & 1 & 16 & 2h30 & 180 \\
          8 pilots & 1 & 56 & 1 & 8 & 2h30 & 180 \\
          16 pilots & 1 & 28 & 1 & 4 & 2h30 & 180 \\

@@

       For batch, we requested 1 to 4 dedicated nodes, depending on resource 
       configuration, each with 112~G of RAM, 1 core per task and 16 tasks per 
       node. On the other hand, for all configurations, we ran our experiments
       with 8 and 16 pilots. Experimental conditions can be seen in Table~\ref{table:conditions}.

@@

       For each configuration, all three execution modes were executed in parallel
       to ensure that the status of the overall cluster was the same when the 
       different execution modes were launched. Furthermore, the order of the execution modes
       for each configuration was randomized to ensure that Slurm job request
       order could not have affected results.

@@

As can be seen in both Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar}, pilots 
generally did not bring any performance improvement to the execution of Spark workloads
on either system. In B\'eluga and Cedar, 16 pilots were generally slower than batch except for a 
few occurrencs. When 16 pilots were faster, it was generally either very slight, or, as expected,
when batch queuing times were very large. The maximum speedup achieved
was approximately 5.6~x for 8 and 16 pilots on B\'eluga and 2.6~x on Cedar. 
It is only in Configuration~3 (Figure~\ref{fig:beluga4})
that we spot any large speedups, however, it generally remains that
pilots are slightly slower, even in Configuration~3. The same can be said about Cedar, however,
it occured with Configuration~4 instead of 3. On average, pilots were up to 0.11~x slower on B\'eluga and
0.17~x slower on Cedar (see Table~\ref{table:speedup}), both of which occurred when using 16 pilots.
The only times when pilots were found to be, on average, faster, was in Configuration~3 on B\'eluga
and Configuration~4 on Cedar where the speedup was found to reach a max of 1.46 for 8 pilots on 
B\'eluga.

@@

%Technical overhead of using pilots + an overlay cluster over a cluster.
Pilots do not appear to provide any performance advantage over batch scheduling.
This can be due to a few reasons. The queuing time for batch and pilots were similar, 
which can be an indication that we were just lacking the right level of priority to 
have been scheduled immediately. In the instances where batch was found to be significantly slower,
it might have been then case that we had obtained the necessary priority level but there
were not enough resources available to be scheduled.

@@

\section{Conclusion}\label{sec:conclusion}
Pilots do not provide significant advantage over batch systems using the same amount of resources. This is affected
by the overhead of dynamically adjusting the cluster size and the fact that queuing time of the batch applications took
as long as the pilots. These results, however, are to an extent cluster-dependent. It is expected that as the number
of resources requested increases, the speedup provided by pilots will increase. Furthermore, users executing on a smaller cluster 
may experience greater speedups using the same resource configuration.

>>>>

\documentclass{IEEEtran}
\usepackage[hyphens]{url}
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{colortbl} % for \rowcolor
\usepackage{ulem} % for \uwave
\usepackage{caption}
\usepackage{subcaption}
\usepackage[bookmarks=false]{hyperref}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=black,citecolor=blue,filecolor=black,urlcolor=blue}
\usepackage{algorithm}
\usepackage{algorithmicx} % Doc is at http://tug.ctan.org/macros/latex/contrib/algorithmicx/algorithmicx.pdf
\usepackage{algpseudocode}
\usepackage{stmaryrd}

\algnewcommand\algorithmicforeach{\textbf{for each}}
\algdef{S}[FOR]{ForEach}[1]{\algorithmicforeach\ #1\ \algorithmicdo}
\algblock{Input}{EndInput}
\algnotext{EndInput}
\algblock{Output}{EndOutput}
\algnotext{EndOutput}
\newcommand{\Desc}[2]{\State \makebox[2em][l]{#1}#2}

@@

\begin{abstract}
    Big Data has become prominent throughout many scientific fields and, as
    a result, scientific communities have sought out Big Data frameworks to
    accelerate the processing of their increasingly data-intensive
    pipelines. However, while scientific communities typically rely on
    High-Performance Computing (HPC) clusters for the parallelization of
    their pipelines, many popular Big Data frameworks such as Hadoop and
    Apache Spark were primarily designed to be executed on dedicated
    commodity infrastructures. This paper evaluates the benefits of pilot
    jobs over traditional batch submission for Apache Spark on HPC
    clusters. Surprisingly, our results show that the speed-up provided by
    pilot jobs over batch scheduling is moderate to inexistent (0.98 on
    average) despite the presence of long queuing times. In addition, pilot
    jobs provide an extra layer of scheduling that complexifies debugging
    and deployment. We conclude that traditional batch scheduling should
    remain the default strategy to deploy Apache Spark applications on
    HPC clusters.

@@

Pilot jobs, also known as dynamic resource provisioning or glide-in
scheduling, are a widely-used technique to address infrastructure
heterogeneity, variable task queuing times, fine task granularity, and node
failures on distributed computing infrastructures. Made popular by software
projects such as Condor~\cite{thain2005distributed} and DIRAC~\cite{casajus2010dirac}, they
became critical to grid computing, greatly improved the performance of HPC
clusters, and enabled multi-cloud executions.

@@

% Goal of the paper
In this paper, we study the use of pilot jobs to run scientific
applications with Apache Spark~\cite{zaharia2016apache} on shared HPC
clusters. The method currently recommended for this purpose involves batch
requesting all the
necessary resources and launching a Standalone Spark cluster once the
resources have been allocated. This is for instance the method used in
Compute Canada,
 our national computing infrastructure. With pilot
jobs, rather than requesting all the resources at once, a Spark cluster is
launched with a subset of the resources, and is expanded as more resources
get allocated. Consistently with previous examples of pilot job
deployments, we hypothesize that this strategy would reduce queuing times
by (1) fragmenting resource requirements, and (2) allowing shorter walltime
estimates. While there have been some efforts on implementing pilot jobs
for Apache Spark~\cite{jha-spark-pbs}, research is limited and
none of the them detail their quantitative effect.

Apache Spark is a popular Big Data framework, commonly used in both
industrial and academic settings. Although it is a Scala-based framework,
it also has APIs for Java, Python (PySpark) and R. Spark's Resilient
Distributed Dataset (RDD) abstraction enabled in-memory processing of
pipelines by co-locating tasks and data, which provided important
performance improvements compared to its predecessor Hadoop
MapReduce~\cite{dean2008mapreduce}. Through the use of RDDs, it also became possible
to execute iterative workflows -- something not easily doable in older
frameworks. Schedulers for Spark include its built-in standalone schedule,
Yet Another Resource Negotiator (YARN~\cite{apache13yet}), and Mesos~\cite{hindman2011mesos}. As a
result of the sustained growth in data volumes, Apache Spark and other
Big Data engines are increasingly used for scientific applications,
including in neuroimaging, our primary field of
interest~\cite{boubela2016big,mehta2017comparative,maybethesimulationone,freeman2014mapping}.

@@

    \subsection{Infrastructure}
    All experiments were conducted on the Cedar and B\'eluga HPC computing
    clusters made available by \href{https://www.computecanada.ca}{Compute
    Canada} through \href{https://www.westgrid.ca}{WestGrid} and
    \href{http://www.calculquebec.ca}{Calcul Qu\'ebec}. Both clusters are
    accessible through the Slurm batch scheduler and Lustre parallel file
    system~\cite{schwan2003lustre}. The Cedar cluster has a total of 1542
    nodes with a total of 58,416 CPU cores. Available memory on a Cedar
    node can range from 125 to 3022~GB. Standard nodes are equipped with
    either 2x Intel E5-2683 v4 Broadwell @ 2.1~Ghz (32 cores total) or 2x
    Intel Platinum 8160F Skylake @ 2.1Ghz (48 cores total) CPUs and 2 x
    480~GB SSD. All nodes and temporary storage on Cedar are connected by an
    Intel OmniPath (version 1) with 100Gbit/s bandwidth.

    B\'eluga, on the other hand, is a smaller cluster with 872 available
    nodes. Node memory can range between 92 to 752~GB, with the most common
    node type having 186G. All nodes contain 2 x Intel Gold 6148 Skylake @
    2.4~Ghz (40 cores/node) CPU and are connected to each other with a
    100~Gb/s Mellanox Infiniband EDR network. Each non-GPU
    node type contains one 480~GB SSD.

    It is important to note that these clusters were used in production,
    that is, concurrently with other users. This allowed us to test the
    added-value of pilot scheduling in different realistic conditions of queuing
    times. At the time of our experiments, B\'eluga had recently been
    commissioned, resulting in low usage and shorter queuing times overall.
    Conversely, Cedar had been operating for a few years, resulting in
    higher usage and longer queuing times. We repeated our experiments multiple 
    times in each configuration to capture queuing time variability.

@@

    A Spark cluster is made up of three components: the cluster manager, workers and
    driver. The cluster manager, also known as the Spark master in Standalone mode, is responsible for the
    resource provisioning within the cluster. Unlike the workers and driver, it is not
    application specific. The driver, which can start in the client where it is launched (i.e. client deploy mode)
    or directly on a worker node (i.e. cluster deploy mode) \TG{Deploy modes are explained in the next paragraph,
    I'd remove them from here}, must request resources from the master
    to launch its tasks. The master dispatches the requests to workers with the necessary resources available to start
    executors. The driver can then communicate with the executors to schedule and launch its tasks. Both the driver and
    the executors are application specific and connected directly to the master.

@@

    There is not shortage of scripts available online for starting a Spark cluster on HPC, 
    see for instance \href{https://github.com/NIH-HPC/spark-slurm}{spark-slurm},
    \href{https://www.sherlock.stanford.edu/docs/software/using/spark}{scripts} provided
     by Stanford's Sherlock cluster, \href{https://sparkhpc.readthedocs.io}{sparkhpc},
     or \href{https://www.osc.edu/~troy/pbstools/man/pbs-spark-submit}{pbs-spark-submit}.
    However, all these scripts start up a cluster within a single batch call. At most, there are two calls, one to startup the cluster,
    and the second, to start the application execution. None of these use any kind of pilot-scheduling approach, nor do they discuss the
    effects of queuing time and how they might want to adjust it in the
    case where the driver is started separately from the batch job to start
    the cluster. Even in the case of RADICAL-Pilot, as per some available
    tutorials for using the pilot-job application with
    Spark (see
    \href{https://github.com/radical-cybertools/pilot-streaming/blob/master/examples/Pilot-Streaming-GettingStarted.ipynb}{here}
    and \href{https://github.com/radical-cybertools/MIDAS-tutorial/blob/master/pilot/Pilot-Spark.ipynb}{here}),
    it appears that the Spark cluster is started within a single pilot and
    the application is submitted, in a separate process, to that pilot. Due
    to the fact that existing solutions mainly rely on a single batch call
    to startup the Spark cluster and do not appear to be able to submit
    multiple smaller jobs to improve queuing times, we have decided to
    implement our own solution for the sake of our experiments.

    Two different job templates were developed to implement the two main
    conditions compared in our experiments: the batch submission
    template and the pilot submission template. The batch submission template was
    inspired by the template provided by Compute Canada to launch Spark applications 
    on Slurm, \href{https://docs.computecanada.ca/wiki/Apache\_Spark/en}{available here}.
    The template operates as follows: certain resource requirements are requested by
    the user (e.g. walltime, amount of memory per node, number of CPUs per task, number 
    of nodes and number of tasks per node). Once these resources are allocated, a 
    master is started on one of the requested node resources. Then, after the master
    has successfully started, the workers are started on all nodes. Multiple worker
    instances are started on a single node by setting the \texttt{SPARK\_WORKER\_INSTANCES}
    environment variable to the number of tasks per node. Each worker is given as many
     cores as specified by the user in the Slurm resource allocation request.
    After both the masters and the workers have successfully started, the driver is finally
    started. The amount of memory given to each executor corresponds to 95\% of the 
    available memory on the node, to allow for offheap space. 
    The Spark deploy mode selected for the batch template is client mode. We selected this
    deploy mode over cluster mode as driver recovery would not be required in batch in the
    case of walltime expiration, as all the allocated resources will expire at the same time.
    Moreover, many scientific applications are written in Python, which cannot use cluster mode using
    the Standalone resource manager. Should pilot scheduling using cluster mode significantly reduce
    queuing time compared to batch scheduling using client mode, it would provide enough justification to extend
    Spark's Standalone scheduler to include cluster mode for PySpark applications.

@@

    Both of these Slurm templates are launched within a Python application called
    \textit{SPA}. The templates are used in conjunction with JSON configuration 
    file and passed to the \href{https://github.com/brentp/slurmpy}{SlurmPy} library within \textit{SPA}.
    The \textit{SPA} application, all the while ensures that all is preconfigured
    correctly before passing it to SlurmPy. It also ensures that enough pilots are
    launched, maintains track of the running/queued pilots, and launches additional
    pilots if there are less pilots than requested by the user in the Slurm queue.

    \subsection{Application}
        \begin{algorithm}\caption{Incrementation}\label{alg:incrementation}              
            \begin{algorithmic}[1]                                                       
                \Input                                                                       
                    \Desc{$x$}{a sleep delay in seconds}                                         
                    \Desc{$n$}{a number of iterations}                                           
                    \Desc{$C$}{a set of image chunks}                                            
                \EndInput                                                                    
                \ForEach{$chunk$ in $C$}                                                      
                    \State read $chunk$                                        
                    \For{$i \in \llbracket 1, n \rrbracket$}                                                         
                        \State $chunk\gets chunk+1$                                              
                        \State sleep $x$                               
                    \EndFor                                                                      
                    \State write $chunk$                                            
                \EndFor                                                                      
            \end{algorithmic}                                                                
        \end{algorithm}
    To determine the added value of pilot scheduling over batch scheduling of Spark
    applications, we required a Spark application operating on a large dataset with
    an important processing time to emulate what would be the average requirements
    of a scientific Spark application. For this, we created a synthetic application 
    that would process the Big Brain~\cite{amunts2013bigbrain}, a 76~GB 3D histological
    image of a human brain. The algorithm is a chain of map transformations that
    at each transformation increment the voxels of the image by 1 (see Algorithm~\ref{alg:incrementation}).
    We chose such a synthetic algorithm
    as the focus of our experiments is pilot scheduling and not the application in
    itself. Furthermore, this algorithm enabled us to have control over the task duration
    which was representative of scientific applications. Additionally, it was important
    that the overall application duration did not vary between the different levels of
    parallelism within our experiments. Being able to adjust the task duration based on
    level of parallelism allowed us to achieve this.

    Investigating the fault-tolerance provided by the overlay cluster is
    crucial in determining the suitability of pilot scheduling with overlay clusters \TG{You should water this down if 
    no fault tolerance experiment is presented}.
    Fault-tolerance of the driver is only possible in cluster deploy mode, however,
    when using Spark's Standalone scheduler, this mode is not available for Python 
    applications. It is for this reason that our synthetic application is written
    in Scala. Nevertheless, cluster deploy mode is possible for Python application
    using YARN or Mesos schedulers.

@@

    Assuming a divisible load, i.e., the application can be divided in any
    number of tasks, the makespan can be written using the following
    expression, which holds for both batch and pilot execution modes:
    \begin{equation}
        M = \frac{C}{W} \label{eq:mcw}
    \end{equation}
    where:
    \begin{itemize}
        \item $M$ is the makespan of the application
        \item $C$ is the total CPU time of the application
        \item $W$ is the \emph{average} number of Spark workers throughout the execution
    \end{itemize}
    The average number of workers $W$ allows the model to take into account
     variable queuing times. It is computed as follows:
    \begin{equation}
        W = \frac{1}{M}\int_0^M{w(t)dt}\label{eq:avgw}
    \end{equation}
    where $w(t)$ is the number of workers available at time $t$. When the
    application is not subject to any scheduling or queuing time, the
    average number of workers equals the number of workers requested.

@@

        \begin{table*}                                                                   
        \centering                                                                       
        \begin{tabular}{c|cccccc}                                                   
          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 1}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          Execution mode & Nodes/job & RAM (GB) & CPUs per task & Tasks/node & Walltime & Task delay (s) \\                             
          \hline
          Batch & 1 & 112 & 1 & 16 & 2h30 & 45 \\
          8 pilots & 1 & 14 & 1 & 2 & 2h30 & 45 \\
          16 pilots & 1 & 7 & 1 & 1 & 2h30 & 45 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 2}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          Execution mode & Nodes/job & RAM (GB) & CPUs per task & Tasks/node & Walltime & Task delay (s) \\                             
          \hline
          Batch & 2 & 112 & 1 & 16 & 2h30 & 90 \\
          8 pilots & 1 & 28 & 1 & 4 & 2h30 & 90 \\
          16 pilots & 1 & 14 & 1 & 2 & 2h30 & 90 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 3}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          Execution mode & Nodes/job & RAM (GB) & CPUs per task & Tasks/node & Walltime & Task delay (s) \\                             
          \hline
          Batch & 3 & 112 & 1 & 16 & 2h30 & 120 \\
          8 pilots & 1 & 42 & 1 & 6 & 2h30 & 120 \\
          16 pilots & 1 & 21 & 1 & 3 & 2h30 & 120 \\

@@

          \rowcolor{headcolor}                                                           
          \multicolumn{7}{c}{Configuration 4}\\                      
          \hline                                                                         
          \rowcolor{headcolor}                                                           
          Execution mode & Nodes/job & RAM (GB) & CPUs per task & Tasks/node & Walltime & Task delay (s) \\                             
          \hline
          Batch & 4 & 112 & 1 & 16 & 2h30 & 180 \\
          8 pilots & 1 & 56 & 1 & 8 & 2h30 & 180 \\
          16 pilots & 1 & 28 & 1 & 4 & 2h30 & 180 \\

@@

       For batch, we requested 1 to 4 dedicated nodes, depending on resource 
       configuration, each with 112~GB of RAM, 1 core per task and 16 tasks per 
       node. On the other hand, for all configurations, we ran our experiments
       with 8 and 16 pilots. Experimental conditions can be seen in Table~\ref{table:conditions}.

@@

       For each configuration, all three execution modes (batch, 8 pilots,
       16 pilots) were executed in parallel to ensure that the status of
       the HPC cluster was the same when the different execution modes
       were launched. Furthermore, the submission order of the execution
       modes for each configuration was randomized to ensure that Slurm job
       request order could not have affected results.

@@

As can be seen in both Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar}, pilots 
generally did not bring any performance improvement to the execution of Spark workloads
on either system. In B\'eluga and Cedar, 16 pilots were generally slower than batch except for a 
few occurrencs. When 16 pilots were faster, it was generally either very slight, or, as expected,
when batch queuing times were very large. The maximum speedup achieved
was approximately 5.6~x for 8 and 16 pilots on B\'eluga and 2.6~x on Cedar. 
It is only in Configuration~3 (Figure~\ref{fig:beluga4})
that we spot any large speedups, however, it generally remains that
pilots are slightly slower, even in Configuration~3. The same can be said about Cedar, however,
it occured with Configuration~4 instead of 3. On average, pilots were up to 0.11~x slower on B\'eluga and
0.17~x slower on Cedar (see Table~\ref{table:speedup}), both of which occurred when using 16 pilots.
The only times when pilots were found to be, on average, faster, was in Configuration~3 on B\'eluga
and Configuration~4 on Cedar where the speedup was found to reach a max of 1.46 for 8 pilots on 
B\'eluga. The missing bars in Figures~\ref{fig:makespansbeluga} and~\ref{fig:makespanscedar} correspond to 
failed executions.

@@

%Technical overhead of using pilots + an overlay cluster over a cluster.
Pilots do not appear to provide any performance advantage over batch
scheduling to deploy Apache Spark overlay clusters on HPC.
This can be due to a few reasons. The queuing time for batch and pilots were similar, 
which can be an indication that we were just lacking the right level of priority to 
have been scheduled immediately. In the instances where batch was found to be significantly slower,
it might have been then case that we had obtained the necessary priority level but there
were not enough resources available to be scheduled.

@@

\section{Conclusion}\label{sec:conclusion}
Pilots do not provide significant advantage over batch systems to run
Apache Spark applications on HPC clusters. This is affected by the overhead of dynamically
adjusting the cluster size and the fact that queuing time of the batch
applications took as long as the pilots. These results, however, are to an
extent cluster-dependent. It is expected that as the number of resources
requested increases, the speedup provided by pilots will increase.
Furthermore, users executing on a smaller cluster may experience greater
speedups using the same resource configuration. Our resource requirements
and cluster size, however, match typical configurations currently available
to scientific communities.
